---
title: "Non standard evaluation examples"
output:
  markdown_document:
    default
---

# Non standard evauation (NSE) intutition and examples

Non-standard evaluation is the evaluation of code in non-standard ways.  Often, this means capturing expressions before they are evaluated, and evaluating them in a different environment (context/scope) to normal.  

## Examples of use of dplyr::select

```{r, echo=TRUE}
df <- tibble::data_frame(a = 1:5, b = 6:10, c = 11:15, d = 16:20)
```

A simple example of the select statement is as follows:
```{r, echo=TRUE}
r <- dplyr::select(df, a, b)
```
This is an example of NSE because a and b are not variables that exist in the global environment.  Instead of searching for a and b in the global namespace, `dplyr::select` directs R to search for the variables a and b in the context of dataframe `df`.  You can think of the environment a bit like a list, and a and b as keys.  So the following is a bit like telling R to lookup `df$a` and `df$b`

Function arguments in R are [promises](https://cran.r-project.org/doc/manuals/r-release/R-lang.html#Promise-objects) which are not evaluated immediately.  They can be captured as expressions and then run in a different environment.

This is fine if we know we want to select the columns `a` and `b` in advance.  But what if these columns are unknown in advance, and are held in a variable.

```{r, echo=TRUE}
columns_to_select <- c("a", "b")
```

The following does not work:

```{r error=TRUE}
dplyr::select(df, columns_to_select)
```
 This error is telling us that there is no column called 'columns_to_select' in the dataframe.  The argument `columns_to_select` has been evaluated in the context of the dataframe, so R has tried to do something like `df$columns_to_select`, and found that the column does not exist.

How do we fix this?

Tidyverse functions always provide an ['escape hatch'](http://adv-r.had.co.nz/Computing-on-the-language.html#substitute) that allow you to get around this limitation.   The [dplyr vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/nse.html) says 'Every function in dplyr that uses NSE also has a version that uses SE. The name of the SE version is always the NSE name with an _ on the end.'

What does this mean?

We might try the following, but we find it does not work:
```{r}
# Does not work
r <-dplyr::select_(df, columns_to_select)
```
As opposed to capturing the argument `columns_to_select` to the select_ function and interpreting it as a column name, `columns_to_select` is evaluated in a standard way, resolving to `c("a", "b")`.

That's what we want, except that each argument to `select_` is a single column, and we've just provided a character vector of length two to represent a single column.

The above code therefore returns a tibble with a single column, a, which is not what we wanted.  (Only the first element - `"a"` in the character vector is used, everything else is ignored).

One solution to this problem is as follows, but it assumes that `columns_to_select` contains exactly two elements:

```{r}
col1 <- columns_to_select[1]
col2 <- columns_to_select[2]
r <- dplyr::select_(df,col1, col2)
```

How do we generalise this to the case where `columns_to_select` may have an arbitrary number of elements?

The solution is to use the optional `.dots` argument.

```{r}
dplyr::select_(df, .dots=columns_to_select)
```

This bears some explanation

In R, the [`...` construct](https://cran.r-project.org/doc/manuals/r-release/R-lang.html#Dot_002ddot_002ddot) allows the creation of functions with a variable (arbitrary) number of arguments.  The `...` is available _within_ the function, and allows the function body to access all of the arguments.  See also [here](https://www.r-bloggers.com/r-three-dots-ellipsis/).

A very simple example is as follows:

```{r}
addition <- function(...) {
  args <- list(...)
  sum(unlist(args))
}
r <- addition(1,2,3)
```

However, this doesn't immediately help us here.  It's actually already implemented in the `select_` function and merely enables us to provide an arbitrary number of column names as arguments, e.g. `select_(df, "a", "b", "c", "d")`.  

What we need is a mechanism that is similar to `...`, but allows us to pass something like `...` into the function as a single argument.  This is exactly what `.dots` does.

Note that `.dots` is not provided by `select`, because this is designed to be used interactively.

## Implementation

The actual implementation of `select` and `select_` is complex.  We can build some heavily simplfied version of these functions, which are not fully functional, but do help to understand what's going on:


```{r}

select2 <- function(df, ...) {
  s <- substitute(list(...))
  eval(s, df) #This evaluates each part of s and returns a list
}

r <- select2(df, a,b)

```

```{r}

select2_ <- function(df, ..., .dots) {
  
  len <- length(list(...))
  
  if (len) {
    lst <- list(...)
  } else {
    lst <- .dots
  }
  
  lst_exp <- lapply(lst, as.name)
  lapply(lst_exp, eval, df) 
}

select2_ <- function(df, ..., dots) {
   len <- length(list(...))
  
  if (len) {
    lst <- list(...)
  } else {
    lst <- .dots
  }
   
   df[unlist(lst)]
}



select2_(df, "a", "b")
select2_(df, .dots = c("a", "b"))



```

## Further reading

[Advanced R material](http://adv-r.had.co.nz/Computing-on-the-language.html)

[Lazyeval vignette](https://cran.r-project.org/web/packages/lazyeval/vignettes/lazyeval.html)

[Dplyr NSE vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/nse.html)