---
title: "Non standard evaluation examples"
output:
  html_document: default
  html_notebook: default
---

# Non standard evauation (NSE) examples

Non-standard evaluation:  the evaluation of code in non-standard ways.  Often, this means capturing expressions and evaluating them in a different environment.  Expressions can be captured because function arguments in r are promises which are not evaluated immediately.

## Reading material
[Advanced R material](http://adv-r.had.co.nz/Computing-on-the-language.html)

[Lazyeval vignette](https://cran.r-project.org/web/packages/lazyeval/vignettes/lazyeval.html)

[Dplyr NSE vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/nse.html)


```{r, echo=TRUE}
df <-tibble::data_frame(a = 1:5, b = 6:10, c = 11:15, d = 16:20)

```
The following works.  This is an example of NSE because a and b are not variables in the global environment (namespace).  Instead of searching for a and b in the global namespace, `dplyr::select` directs R to search for the variables a and b in the context of dataframe `df`.  You can think of the environment a bit like a list, and a and b as keys.  So the following is a bit like `df$a` and `df$b`

```{r, echo=TRUE}
r <- dplyr::select(df, a, b)
```

```{r, echo=TRUE}

# Does not work - each one of the arguments to select must be a single column
r <- dplyr::select_(df, c("a", "b"))  

#Works - this is the 'standard evaluation' version.
r <- dplyr::select_(df, "a", "b")
#What does this do?  it takes "a" and "b" and evaluates them within the context of df

#Works 
r <- dplyr::select_(df, .dots=c("a", "b"))


#Works
col1 = "a"
r <- dplyr::select_(df, col1)
#This has taken the variable col1 and evaluted it in the contexxt of df

```
```{r}
#We could start to think about implementing a selection function like this:


select2 <- function(df, ...) {
  s <- substitute(list(...))
  print(deparse(s))
  eval(s, df) #This evaluates each part of s and returns a list
}


r <- select2(df, a,b)
r <- select2(df, a=a,b=b) 

#Note even this will work

r <- select2(df, a=sin(a), b=b)

```

```{r}
#Possible implementation of select_

#Why does select_ need ... and .dots when select uses ... only
select2_ <- function(df, .dots, ...) {
  if (length(list(...))) {
    eval(list(...), df)
  } else {
    eval(.dots, df)
  }
}

select2_(df, a="a", b="b")
select2_(df, .dots=list(a="a", b="b"))


```

