---
title: "Non standard evaluation tools"
author: "Robin Linacre"
date: "11 June 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pryr)
library(lazyeval)

get_str <- function(x) {
  expr_str <- deparse(substitute(x))
  expr_str <- stringr::str_replace(expr_str,  "capture\\.output\\((.{1,20})\\)", "\\1") 
  stringr::str_interp("$[-17s]{expr_str} : ${trimws(x)}")
}

get_info <- function(fn_input) {
  print(stringr::str_interp("fn_input is ${deparse(substitute(fn_input))}."))
  print(get_str(capture.output(print(fn_input))))
  print(get_str(typeof(fn_input)))
  print(get_str(class(fn_input)))
  print(get_str(capture.output(str(fn_input))))
}

get_str2 <- function(expr, fn_to_apply) {
    expr <- lazyeval::expr_find(expr)
    expr_c <- deparse(substitute(fn_to_apply(expr)))
    value <- fn_to_apply(eval(expr))
    stringr::str_interp("$[-17s]{expr_c} : ${trimws(value)}")
}

get_info2 <- function(expr) {
  print(get_str2(expr, typeof))
  print(get_str2(expr, class))
  print(get_str2(expr, print))
  print(get_str2(expr, str))
  
}
```


# Non-standard evaluation - tools

## Base R

### Quote

Quote returns its argument as an unevaluated expression `expr`
```{r}
x <- 1
get_info(quote(x))
get_info(x)
get_info(1)
```


### Substitute

`substitute` returns the parse tree for the (unevaluated) expression expr, **substituting any variables bound in env.**

Unfortunately substitute() has a feature that makes modifying calls interactively a bit of a pain. When run from the global environment, it never does substitutions: in fact, in this situation it behaves just like quote().




```{r}
x <- 1
get_info(substitute(x))
get_info(substitute(x+y))

f <- function() {
  x <- 1
  get_info(substitute(x+y))
}
f()

```

When used within a function, substitute is able to capture the arguments to the function.  It looks at a function argument and instead of seeing the value, it sees the code used to compute the value.

```{r}
f <- function(arg) {
  substitute(arg)
}
get_info(f(abc))

```




### as.name

So far, we've seen ways of manipulating expressions, including resolving parts of the expression.

as.name does the opposite.  It is a way of taking a string and making it into an expression
```{r}
col <- "mpg"
get_info(as.name(col))
get_info(as.name("mpg"))
get_info(quote(col))

mpg <- 10
eval(as.name(col))

```

### parse

Parse generalises this idea.  It returns an expression which is a group of symbols

```{r}
a <- 1
b <- 2
t <- "a + b"

get_info(parse(text=t))
get_info(parse(text="a + b"))

eval(parse(text = t))
eval(parse(text = "a + b"))

get_info(parse(text="a"))
```

### ~ (the symbol to create a formula)

Technically, a formula is a “language” object (i.e. an unevaluated expression) with a class of “formula” and an attribute that stores the environment.

~ is a single character that allows you to say: “I want to capture the meaning of this code, without evaluating it right away”. For that reason, the formula can be thought of as a “quoting” operator.


```{r}
get_info(~x)
```

You can access the expression in the formula like this:

```{r}
x <-1
f <- ~x
get_info(f[[2]])
get_info(eval(f[[2]]))
```

The difference between these two is that in the first one, as soon as x is used, it resolves to its value.  In the second one, the arg is an expression that has the potential to be evaluted, but is only actually evaluated when you tell it to 
```{r}
x <- 1

f <- function(arg) {
  arg
}

f(x)
f(~x)
eval(f(~x)[[2]])
```

## How to use these tools (lazyeval and tidyverse)

Lazyeval [is](https://cran.r-project.org/web/packages/lazyeval/vignettes/lazyeval.html) a package that provides principled tools to perform non-standard evaluation (NSE) in R. 

It recommends the use of formulas for non standard evaluation.  Formulas provide a way of 'quoting' code (converting it into an expression which is not evaluated until asked), whilst retaining the environment it was created it.

This allows you to use a formula as a robust way of delaying evaluation, cleanly separating the creation of the formula from its evaluation.

Lazyeval provides a bunch of tools that make it easier to work with delayed evaluation - providing alternative to the base R function described above.

### Example 1 - dplyr aggregate 

Supposing 

```{r}

cols <- c("a", "b")




```

